---
mydisqus: "public_API"
mytitle: "Cryptocurrency Price Data Collection"
mysub: "Comprehensive Market Data Harvesting: 800,000+ Data Points Across 5 Exchanges"
output:
  html_document:
      self_contained: false    
      mathjax: default
      lib_dir: libs
      template: template_crypto.html
      theme: "lumen"
      css: style.css
      toc: FALSE
      toc_float: TRUE
      toc_depth: 2
      number_sections: TRUE
      df_print: "paged"
      code_folding: "hide"
      includes: 
        after_body: footer.html
---


<div class="mycontent">

# Public Exchange API Integration
***

Cryptocurrency markets operate through specialized trading platforms called **exchanges**, which function as digital marketplaces for buying and selling digital assets. Unlike traditional financial institutions, these exchanges operate 24/7 with prices fluctuating continuously based on real-time supply and demand dynamics, often updating every few seconds.

<br>

Modern cryptocurrency exchanges provide **public APIs** (Application Programming Interfaces) that enable programmatic access to real-time market data. For instance, to retrieve the current Bitcoin price from the Kraken exchange, you can access the following endpoint directly in your browser:

[https://api.kraken.com/0/public/Ticker?pair=BTCEUR](https://api.kraken.com/0/public/Ticker?pair=BTCEUR)

This API call returns comprehensive market data including three critical price points:

- **`last`**: The price of the most recent completed transaction
- **`bid`**: The highest price buyers are willing to pay (sell price for holders)
- **`ask`**: The lowest price sellers are willing to accept (buy price for purchasers)

# Programmatic Data Collection
***

The real power of public APIs lies in their ability to be accessed **programmatically**, enabling automated data collection at scale. This capability is essential for systematic market analysis and algorithmic trading strategies.

Below is a practical example using the R programming language to retrieve Bitcoin market data from Kraken and format it into a structured table:

```{r, warning=FALSE, message=FALSE, eval=TRUE}
# Load required packages
library(tidyverse)

# Establish connection and retrieve market data
library(RCurl)
api_endpoint <- "https://api.kraken.com/0/public/Ticker?pair=BTCEUR"
raw_data <- getURLContent(api_endpoint)

# Parse JSON response and structure data
require(jsonlite)
parsed_data <- fromJSON(raw_data)$result[[1]]
market_data <- data.frame(
  ask = parsed_data$a[1], 
  bid = parsed_data$b[1], 
  last = parsed_data$c[1], 
  open = parsed_data$o, 
  low = parsed_data$l[1], 
  high = parsed_data$h[1], 
  volume = parsed_data$v[1], 
  volumeQuote = NA, 
  timestamp = NA
)

# Display results
```

```{r, warning=FALSE, message=FALSE, eval=TRUE, echo=FALSE}
# Format and display market data
library(kableExtra)
options(knitr.table.format = "html")

formatted_data <- market_data %>% 
  select(-volumeQuote, -timestamp)
apply(formatted_data, 2, function(x){round(as.numeric(x),0)}) %>%
  t() %>%
  kable() %>% 
  kable_styling(bootstrap_options = "striped", full_width = F)
```

# Comprehensive Function Library
***

To facilitate systematic data collection across multiple exchanges, I have developed a comprehensive [function library](https://github.com/ymju86/Crypto-Arbitrage/blob/master/FUNCTIONS/Public_Market_Functions.R) that standardizes API interactions across five major cryptocurrency exchanges. This library provides unified access to market data regardless of the underlying exchange's API structure.

**Usage Example:**
```{r, warning=FALSE, message=FALSE, eval=FALSE}
# Load the standardized API functions from GitHub repository
source("https://raw.githubusercontent.com/ymju86/Crypto-Arbitrage/master/FUNCTIONS/Public_Market_Functions.R")

# Retrieve Bitcoin price from Bitstamp exchange
get_bitstamp(Sys.time(), "BTCEUR")
```

# Large-Scale Data Harvesting Operation
***

This research project involved an extensive **data collection campaign** with the following specifications:

- **Duration**: 14-day continuous monitoring period (February 5-19)
- **Cryptocurrencies**: 5 major digital assets - Bitcoin (BTC), Bitcoin Cash (BCH), Ethereum (ETH), Litecoin (LTC), and Ripple (XRP)
- **Exchanges**: 5 leading platforms - Coinbase, Kraken, Bitstamp, Bitfinex, and CEX.io
- **Data Points**: Over 800,000 individual price observations
- **Frequency**: Real-time data collection every few seconds

The data harvesting was implemented using an **automated collection system** that continuously called the API functions described above. The complete implementation script is available [here](https://github.com/ymju86/Crypto-Arbitrage/blob/master/FUNCTIONS/Run_API_public_market_harvest.R).

<br>

**Data Access:**
The complete dataset is publicly available on GitHub in compressed format. You can load and analyze the data using the following R code:

```{r, warning=FALSE, message=FALSE, eval=FALSE}
# Load the comprehensive market dataset
load("../DATA/public_ticker_harvest.Rdata")  

# Examine the dataset structure
head(Ticker)
```

<br>

**Data Visualization Preview:**
As a demonstration of the dataset's richness, here's the Ethereum price evolution on Bitstamp during the collection period:

```{r, fig.align="center", warning=FALSE, cache=TRUE, fig.width=9, message=FALSE, fig.height=4}
# Load the comprehensive dataset
load("../DATA/public_ticker_harvest.Rdata")

# Create price evolution visualization
library(hrbrthemes)
Ticker %>%
  filter( symbol == "ETHEUR" ) %>%
  filter(platform == "Bitstamp") %>%
  ggplot( aes(x=time, y=as.numeric(last))) +
    geom_ribbon(aes(ymin=450, ymax=as.numeric(last)),  fill="#69b3a2", color="transparent", alpha=0.5) +
    geom_line(color="#69b3a2") +
    ggtitle("Ethereum Price Evolution During Data Collection Period") +
    ylab("Ethereum Price (EUR)") +
    theme_ipsum() +
    theme(
      plot.title = element_text(size=12)
    )
```

# Next Phase: Market Analysis
***

With this comprehensive dataset in hand, the next phase involves **quantitative analysis** of price differences between exchanges. If these differences prove significant enough, they may present viable arbitrage opportunities.

<br>
<center>
<a class="btn btn-primary" href="https://ymju86.github.io/Crypto-Arbitrage/CONTENT/differences.html">Analyze Price Differences →</a>
</center>
<br><br>

</div>













